---
title: "Ridgeline Flux Plots"
output:
  html_document:
    df_print: paged
---

The dataset has the following columns:

- `Caldera`: This specifies the caldera name.
- `Site_Name`: The name of the site within the caldera.
- `Soil_Type`: The type of soil at the site.
- `Date_of_Measurement`: The date when the measurement was made.
- `CH4_Flux`: The methane flux.
- `log10FCH4`: The base-10 logarithm of the methane flux.
- `d13CH4_source`: The 㭁3C value for the methane source.
- `CO2_Flux`: The carbon dioxide flux.
- `log10FCO2`: The base-10 logarithm of the carbon dioxide flux.
- `d13CO2_source`: The 㭁3C value for the carbon dioxide source.

Below is R code (using **tidyverse**, **ggplot2**, and **ggridges**) to create **ridgeline density plots** of `log10FCO2` and `log10FCH4` for each `Site_Name`, **stacked vertically** with a shared x-axis in each plot. Ridgelines are **grouped by `Caldera` then `Soil_Type`** via the y-axis ordering, and **colored by `Soil_Type`**. Both plots use the **same site order** and include **horizontal guideline separators** between sites. No plot title is used. Legends are placed at the top and forced onto one row.

# Load packages

```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)
```

# ---- 1) Read and prepare data ----

```{r}
df <- readr::read_csv("BothCalderasFluxesIsotopes.csv",
                      show_col_types = FALSE)

# Keep only the variables needed for plotting and ensure consistent types
plot_df <- df %>%
  transmute(
    Caldera = as.factor(Caldera),
    Site_Name = as.factor(Site_Name),
    Soil_Type = as.factor(Soil_Type),
    log10FCO2 = as.numeric(log10FCO2),
    log10FCH4 = as.numeric(log10FCH4)
  )

# Create a single, consistent vertical ordering for sites:
# grouped by Caldera, then Soil_Type, then Site_Name
site_order <- plot_df %>%
   distinct(Caldera, Soil_Type, Site_Name) %>%
  arrange(Caldera, Soil_Type, Site_Name) %>%
  mutate(Site_Name = as.character(Site_Name)) %>%
  pull(Site_Name)

plot_df <- plot_df %>%
  mutate(
    Site_Name = factor(as.character(Site_Name), levels = site_order)
  )

# Y positions for guideline separators (one per site level)
guidelines <- tibble(Site_Name = levels(plot_df$Site_Name))
```

# ---- 2) Common theme settings ----

```{r}
ridge_theme <- theme_minimal(base_size = 12) +
  theme(
    plot.title = element_blank(), # explicit: no title
    legend.position = "top",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_blank(),
    # Force legend onto one line when possible
    legend.text = element_text(size = 10),
    panel.grid.major.y = element_blank(), # we'll draw our own site separators
    panel.grid.minor = element_blank()
  )

```

# ---- 3) CO2 ridgeline plot ----

```{r}
p_co2 <- ggplot(
  plot_df %>% filter(is.finite(log10FCO2)),
  aes(x = log10FCO2, y = Site_Name, fill = Soil_Type)
) +
  # Guidelines separating sites
  geom_hline(
    data = guidelines,
    aes(yintercept = 1:length(site_order)),
    inherit.aes = FALSE,
    color = "grey85",
    linewidth = 0.4
  ) +
  # Kernel density ridgelines
  geom_density_ridges(
    scale = 1.2,
    alpha = 0.8,
    color = "grey30",
    size = 0.2,
    rel_min_height = 0.01
  ) +
  coord_cartesian(xlim = c(0, 5)) +
  scale_x_continuous(limits = c(0, 5)) +
  guides(fill = guide_legend(nrow = 1, byrow = TRUE)) +
  labs(x = expression(log[10](F[CO[2]])), y = NULL) +
  ridge_theme
print(p_co2)
```

# ---- 4) CH4 ridgeline plot ----

```{r}
# The filter step below changes the number of riddges, so I need to
# recalculate the number of ridges
ch4plot_df <- filter(is.finite(log10FCH4))
p_ch4 <- ggplot(
  plot_df %>% filter(is.finite(log10FCH4)),
  aes(x = log10FCH4, y = Site_Name, fill = Soil_Type)
) +
  # Guidelines separating sites
  geom_hline(
    data = guidelines,
    aes(yintercept = 1:length(site_order)),
    inherit.aes = FALSE,
    color = "grey85",
    linewidth = 0.4
  ) +
  # Kernel density ridgelines
  geom_density_ridges(
    scale = 1.2,
    alpha = 0.8,
    color = "grey30",
    size = 0.2,
    rel_min_height = 0.01
  ) +
  coord_cartesian(xlim = c(-4, 2)) +
  scale_x_continuous(limits = c(-4, 2)) +
  guides(fill = guide_legend(nrow = 1, byrow = TRUE)) +
  labs(x = expression(log[10](F[CH[4]])), y = NULL) +
  ridge_theme
print(p_ch4)

```

# ---- 5) Export figures (PNG for preview, SVG for publication) ----

```{r}
# CO2
ggsave("ridgeline_log10FCO2.png", plot = p_co2, width = 9, height = 10, dpi = 300)
ggsave("ridgeline_log10FCO2.svg", plot = p_co2, width = 9, height = 10)

# CH4
ggsave("ridgeline_log10FCH4.png", plot = p_ch4, width = 9, height = 10, dpi = 300)
ggsave("ridgeline_log10FCH4.svg", plot = p_ch4, width = 9, height = 10)
```
```
## Output files created

- **`ridgeline_log10FCO2.png`**: Preview-quality PNG of the CO₂ ridgeline density plot (`log10FCO2`) with x-axis limited to **-2 to 5**, legend at the top (one row), and horizontal site-separating guidelines.
- **`ridgeline_log10FCO2.svg`**: Publication-quality SVG version of the same CO₂ plot (vector format).
- **`ridgeline_log10FCH4.png`**: Preview-quality PNG of the CH₄ ridgeline density plot (`log10FCH4`) with x-axis limited to **-4 to 4**, using the **same site ordering** and styling as the CO₂ plot.
- **`ridgeline_log10FCH4.svg`**: Publication-quality SVG version of the same CH₄ plot (vector format).
